Everything you wanted to know about PanDA but were truly truly afraid to ask

### 2017-03-06 PandaEvolution.tex
Small editing
Eliminate event generation

Currently, no HEP application can benefit from Titan's GPUs but some
computationally-intensive and non memory-intensive tasks can be offloaded from
the Grid to Titan's large amount of cores. Further, when HEP tasks can be
partitioned to independent jobs, Titan work nodes can be used to execute up to
16 concurrent jobs' payload, one for each available core. Given these
constraints and challenges, the types of task most suitable for execution on
Titan is Monte Carlo event generation and detector simulation. Both types of
task are mostly computational and stand-alone code, requiring less than 1GB of
RAM at runtime and with small input data requirements. The event generation in
ATLAS accounts for 10--15\% of all jobs on WLCG, corresponding to 2/3 of the
current Grid capacity. Detector simulation accounts for \ldots .

\mtnote{Currently, is PanDA running event generation, detector simulation or
both on Titain?}
\mtnote{Please provide a diagram illustrating event generation and detector
simulation stats compared to the whole ATLAS computation activity.}

Event generation and detector simulation are part of the ATLAS production Monte
Carlo (MC) workflow (also known as MC production chain)~\cite{mcworkflow}. The
MC workflow consists of four main stages: event generation, detector simulation,
digitization, and reconstruction. Even generation creates sets of particle
four-momenta via different generators, e.g., PYTHIA~\cite{herwig} and
HERWIG~\cite{herwig}. The detector simulator is called Geant4~\cite{geant4} and
simulates the interaction of these particles with the sensitive material of the
ATLAS detector. Each interaction creates a so-called hit and all hits are
collected and passed on for digitalization where hits are further process to
mimic the readout of the detector. Finally, reconstruction operates local
pattern recognition, creating high-level objects like particles and jets.

@@

The unavailability of pilots is a potential drawback when executing distributed
workloads like MC event generation and detector simulation. Pilots are used to
increase the throughput of distributed workloads: while pilots have to wait in
the supercomputer's queue, once scheduled, they can pull and execute jobs
independently from the system's queue. This is particularly relevant for
machines like Titan where queue policies privilege parallel jobs on the base of
the number of work nodes they request: the higher the number of nodes, the
shorter the amount of queue time (modulo fair share policies).

@@

\mtnote{Please feel free to add details about backfill as needed.}

@@

Once deployed, every PanDA Broker supports the execution of MC event generation
and detector simulation on Titan in 9 steps: (i) PanDA Broker queries Titan's
Moab scheduler about currently unused worker nodes;
% and checks whether the time and size of the available backfill resources are
% suitable to execute jobs;
(ii) the PanDA Broker queries the Job Dispatcher module of the PanDA server for
jobs that have been bound to Titan by JEDI; (iii)  the PanDA Broker requests to
the Data Service module of the PanDA Server what input data these jobs require
and transfer them to the DTN via the ATLAS DDM service; (iv) the PanDA Broker
creates one or more PBS MPI script, wrapping enough jobs' payload to fit
backfill availability; (v) the PanDA Broker submits these MPI jobs to the
Titan's PBS batch system via RADICAL-SAGA; (vi) the MPI jobs waits a minimal
amount of time into the queue until they are scheduled for execution; (vii) upon
execution on the worker nodes, each MPI script creates a ramdisk and sets up the
specific AthenaMP configuration required by the wrapped payload; (viii) the
payload is executed into dedicated processes, 16 for each work node, 1 for each
available core; (ix) upon completion of an MPI job, the PanDA Broker locates the
produced output files on the shared filesystem and transfer them to BNL ATLAS
Tier 1 computing centre.

>>>>

Currently, no HEP application can benefit from Titan's GPUs but some
computationally-intensive and non memory-intensive tasks can be offloaded from
the Grid to Titan's large amount of cores. Further, when HEP tasks can be
partitioned to independent jobs, Titan work nodes can be used to execute up to
16 concurrent jobs' payload, one for each available core. Given these
constraints and challenges, the type of task most suitable for execution at the moment on Titan is Monte Carlo
% event generation and
detector simulation.
% Both types of task are
This type of task is mostly computational-intensive, requiring less than 1GB of
RAM at runtime and with small input data requirements. Detector simulation tasks
in ATLAS account for XX--XX\% of all jobs on WLCG, corresponding to X/X of the
current Grid capacity.\mtnote{Please provide numbers to replace `XX' and,
possibly, a diagram illustrating detector simulation stats compared to the whole
ATLAS computation activity.}

% The event generation in ATLAS accounts for 10--15\% of all jobs on WLCG,
% corresponding to 2/3 of the current Grid capacity.

% Event generation and
Detector simulation is part of the ATLAS production Monte Carlo (MC) workflow
(also known as MC production chain)~\cite{mcworkflow}. The MC workflow consists
of four main stages: event generation, detector simulation, digitization, and
reconstruction. Even generation creates sets of particle four-momenta via
different generators, e.g., PYTHIA~\cite{herwig} and HERWIG~\cite{herwig}. The
detector simulator is called Geant4~\cite{geant4} and simulates the interaction
of these particles with the sensitive material of the ATLAS detector. Each
interaction creates a so-called hit and all hits are collected and passed on for
digitalization where hits are further process to mimic the readout of the
detector. Finally, reconstruction operates local pattern recognition, creating
high-level objects like particles and jets.

@@

The unavailability of pilots is a potential drawback when executing distributed
workloads like MC
% event generation and
detector simulation. Pilots are used to increase the throughput of distributed
workloads: while pilots have to wait in the supercomputer's queue, once
scheduled, they can pull and execute jobs independently from the system's queue.
Jobs can be concurrently on every core available to the pilot and multiple
generations of concurrent executions can be performed until the pilot's walltime
is exhausted. This is particularly relevant for machines like Titan where queue
policies privilege parallel jobs on the base of the number of work nodes they
request: the higher the number of nodes, the shorter the amount of queue time
(modulo fair share policies).

@@

\mtnote{Please feel free to add details about backfill as needed.}
\mtnote{Please note: I know the following departs for the current name given to
the PanDA subsystem installed on Titan. I think there is a good reason to
call it a broker instead of a pilot, and I think I explained it in the previous
paragraphs. Please  take this just as a suggestion, something I would like to
discuss in our meeting.}

@@

Once deployed, every PanDA Broker supports the execution of MC
% event generation and
detector simulation on Titan in 9 steps: (i) PanDA Broker queries Titan's
Moab scheduler about currently unused worker nodes;
% and checks whether the time and size of the available backfill resources are
% suitable to execute jobs;
(ii) the PanDA Broker queries the Job Dispatcher module of the PanDA server for
jobs that have been bound to Titan by JEDI; (iii)  the PanDA Broker requests to
the Data Service module of the PanDA Server what input data these jobs require
and transfer them to the DTN via the ATLAS DDM service; (iv) the PanDA Broker
creates one or more PBS MPI script, wrapping enough jobs' payload to fit
backfill availability; (v) the PanDA Broker submits these MPI jobs to the
Titan's PBS batch system via RADICAL-SAGA; (vi) the MPI jobs waits a minimal
amount of time into the queue until they are scheduled for execution; (vii) upon
execution on the worker nodes, each MPI script creates a ramdisk and sets up the
specific AthenaMP configuration required by the wrapped payload; (viii) the
payload is executed into dedicated processes, 16 for each work node, 1 for each
available core; (ix) upon completion of an MPI job, the PanDA Broker locates the
produced output files on the shared filesystem and transfer them to BNL ATLAS
Tier 1 computing centre.
